---
title: "FMLM_addmixed"
author: "Kenny Chandra Wijaya"
date: "`r Sys.Date()`"
header-includes:
  - \usepackage{fontspec} # use fontspec package
  - \usepackage{xeCJK}    # use xeCJK package
  - \usepackage{setspace}\setstretch{1.2}
  - \usepackage{titlesec}
  # - \newcommand{\sectionbreak}{\clearpage}
  - \usepackage{siunitx}
  - \newcolumntype{d}{S[table-format=3.3]}
output: 
  pdf_document: 
    pandoc_args: ["--lua-filter=color-text.lua"]
    latex_engine: xelatex
    number_sections: yes
---

```{r setup, include=FALSE}
# Overal all chunk settings

knitr::opts_chunk$set(echo = FALSE,knitr.kable.NA = '')

# I/O
library(readxl)
library(openxlsx)
library(writexl)
# rmd render
library(kableExtra)
library(pander)
library(extrafont)
library(showtext)
# Data cleaning
library(tidyverse)
# Plotting
library(ggpubr)
library(ggplot2)
library(gridExtra)
# DCM
library(apollo)
```

# Preparations

```{r Initialization}

# clean the environment

rm(list = ls())

# Use only SP data
# "C:\Users\Tongfei
# fmlm_5.csv
# fmlm_por.csv
# "C:/Users/24479/Box/[L1FR] STSRG Lab Folder/Prasanna/Papers/FMLM 3-05-24/Purdue/fmlm_por.csv"

database<- read.csv("C:/Users/kenny/Box/[L1FR] STSRG Lab Folder/Prasanna/Papers/FMLM 3-05-24/Purdue/fmlm_5.csv")

database <- database %>% 
  mutate(Pri_car = ifelse(as.numeric(Q3.1_2)==1,1,0)) %>% 
  mutate(Pri_tra = ifelse(as.numeric(Q3.1_2) %in% c(4,5),1,0))
database <- database %>% dplyr::filter(av_car==0)

# glimpse(database)
```

# Apollo settings

## Global control and choice-settings

```{r Apollo_preparation}

# apollo Initialization

model_name = 'addmixed4'
Drawmethod = "sobol"
Ndraws = 2000
apollo_initialise()

# Set core controls

apollo_control = list(
  modelName       = model_name,
  modelDescr      = "MIXL with balanced sociodemo + latent, no correlations",
  indivID         = "cid", 
  outputDirectory = "output",
  nCores          = parallel::detectCores()-1,
  seed            = 123
)

# Define choice and data

choiceAnalysis_settings <- list(
  alternatives = c(walk=1,scooty=2, expbike=3, ondemand=4, ridehail=5),
  avail        = list(walk=database$av_walk, scooty= database$av_Sc, expbike=database$av_Bik,
                      ondemand=database$av_OnD,ridehail=database$av_Rid),
  choiceVar    = database$choice,
  explanators  = database[,c("STUD","Gend")],
  rows = database$choice!=0
)

```

## Define parameters


```{r parameters}

### Vector of parameters

apollo_beta = c(
  # ASC (Alternative Specific Constants)
  asc_scooty         = 0,
  asc_expbike        = 0,
  asc_ondemand       = 0,
  asc_ridehail       = 0,
  
  # Travel time coefficients (mu and sigma)
  b_tt_walk_mu       = -1,
  b_tt_scooty_mu     = -1,
  b_tt_expbike_mu    = -1,
  b_tt_OnD_mu        = -1,
  b_tt_Rid_mu        = -1,
  
  b_tt_walk_sig      = 1,
  b_tt_scooty_sig    = 1,
  b_tt_expbike_sig   = 1,
  b_tt_OnD_sig       = 1,
  b_tt_Rid_sig       = 1,

  # Access time coefficients (mu and sigma)
  b_at_scooty_mu     = -1,
  b_at_expbike_mu    = -1,
  b_at_OnD_mu        = -1,
  b_at_Rid_mu        = -1,
  
  b_at_scooty_sig    = 1,
  b_at_expbike_sig   = 1,
  b_at_OnD_sig       = 1,
  b_at_Rid_sig       = 1,

  # Cost coefficient
  b_cost             = -1,

  # Travel Time (TT) indirect effects
  b_stud_TT_ondemand    = 0,
  b_stud_TT_ridehail    = 0,

  b_scfreq_TT_scooty    = 0,

  b_ridfreq_TT_ridehail = 0,

  # Access Time (AT) indirect effects

  b_stud_AT_ondemand    = 0,
  b_stud_AT_ridehail    = 0,
  
  b_bikfreq_AT_expbike  = 0,
  
  # STUD coefficients for on-demand ride and ride-hailing modes
  b_stud_ondemand    = 0,

  # Latent variable coefficient
  b_GR_ridehail      = 0,  
  
  # Specific individual characteristics
  b_rider_ondemand   = 0,
  b_rider_ridehail   = 0,

  b_scfreq_scooty    = 0,

  b_scbik_plane_scooty  = 0,
  b_scbik_plane_expbike = 0
  )



apollo_fixed = c()

### Set parameters for generating draws

apollo_draws = list(
  interDrawsType = Drawmethod,
  interNDraws = Ndraws,
  interNormDraws= c(# Travel time
                    "draws_tt_sc",
                    "draws_tt_OnD",
                    "draws_tt_bik",
                    "draws_tt_Rid",
                    "draws_tt_walk",
                    "draws_tt_car",
                    # Access time
                    "draws_at_sc","draws_at_OnD",
                    "draws_at_bik","draws_at_Rid")
)

### Create random parameters

apollo_randCoeff = function(apollo_beta, apollo_inputs){
  randcoeff = list()
  
  randcoeff[["b_tt_walk"]] = b_tt_walk_mu + b_tt_walk_sig * draws_tt_walk
  
  randcoeff[["b_tt_sc"]]   = b_tt_scooty_mu + b_tt_scooty_sig * draws_tt_sc  
    
  randcoeff[["b_tt_bik"]]  = b_tt_expbike_mu + b_tt_expbike_sig * draws_tt_bik
                            
  randcoeff[["b_tt_OnD"]]  = b_tt_OnD_mu + b_tt_OnD_sig * draws_tt_OnD
                             
  randcoeff[["b_tt_Rid"]]  = b_tt_Rid_mu + b_tt_Rid_sig * draws_tt_Rid
                            
  randcoeff[["b_at_sc"]]   = b_at_scooty_mu + b_at_scooty_sig * draws_at_sc
                             
  randcoeff[["b_at_bik"]]  = b_at_expbike_mu + b_at_expbike_sig * draws_at_bik
  
  randcoeff[["b_at_OnD"]]  = b_at_OnD_mu + b_at_OnD_sig * draws_at_OnD
     
  randcoeff[["b_at_Rid"]]  = b_at_Rid_mu + b_at_Rid_sig * draws_at_Rid
  
  return(randcoeff)
}

### Read in starting values for at least some parameters from existing 
# apollo_beta = apollo_readBeta(apollo_beta, apollo_fixed, 
#                               "MMNL_heterogeneity_w_ocar+soclat_final1",
#                               overwriteFixed=FALSE)

apollo_inputs = apollo_validateInputs()

```



## Define utility_function

Sociodemographic Controls:
- **Gend**: Gender
- **Emp**: Employment status
- **STUD**: Student status
- **Pri_car**: Private car ownership status
- **Pola**: Political inclination
- **GR**

- **scfreq**: Frequency of scooter usage
- **bikfreq**: Frequency of electric bike usage
- **scbik_plane**: Joint planning variable for electric bike and scooter
- **ridfreq**: Frequency of ride-hailing usage
- **rider**: Whether the person is a rider


```{R utility_function}

apollo_probabilities=function(apollo_beta, apollo_inputs, functionality="estimate"){
  
  ### Attach inputs and detach after function exit
  apollo_attach(apollo_beta, apollo_inputs)
  on.exit(apollo_detach(apollo_beta, apollo_inputs))
  
  ### Create list of probabilities P
  P = list()
  
  ### List of utilities: these must use the same names as in mnl_settings, order is irrelevant
  
  V = list()
  
  # Walk is the base alternative, no individual characteristics included
  V[["walk"]] = (
    b_tt_walk * Walk_TT
  )
  
  # Utility function for electric scooter mode
  V[["scooty"]] = (
    asc_scooty + 
    b_tt_sc * Sc_TT + 
    b_at_sc * Sc_AT + 
    b_cost * Sc_Cost +
      
    # Specific individual characteristics
    b_scfreq_scooty * scfreq +     
    b_scbik_plane_scooty * scbik_plane +
    
    # TT (Travel Time) indirect effects
    (
     b_scfreq_TT_scooty * scfreq 
     ) * Sc_TT 
  
    # AT (Access Time) indirect effects

  )
  
  # Utility function for bike-sharing mode
  V[["expbike"]] = (
    asc_expbike + 
    b_tt_bik * Bik_TT + 
    b_at_bik * Bik_AT + 
    b_cost * Bik_Cost +
  
    # Specific individual characteristics
    b_scbik_plane_expbike * scbik_plane +
    
    # TT (Travel Time) indirect effects
  
    # AT (Access Time) indirect effects
    (
     b_bikfreq_AT_expbike * bikfreq
     ) * Bik_AT
  )
  
  # Utility function for on-demand ride mode
  V[["ondemand"]] = (
    asc_ondemand + 
    b_tt_OnD * OnD_TT + 
    b_at_OnD * OnD_AT + 
    b_cost * OnD_Cost +
  
    # General individual characteristics (GIC)
    b_stud_ondemand * STUD +       

    # Specific individual characteristics
    b_rider_ondemand * rider +      
      
    # TT (Travel Time) indirect effects
    (
     b_stud_TT_ondemand * STUD 
     ) * OnD_TT +
  
    # AT (Access Time) indirect effects
    (
     b_stud_AT_ondemand * STUD 
     ) * OnD_AT
  )
  
  # Utility function for ride-hailing mode
  V[["ridehail"]] = (
    asc_ridehail + 
    b_tt_Rid * Rid_TT + 
    b_at_Rid * Rid_AT + 
    b_cost * Rid_Cost +
  
    # General individual characteristics (GIC)
    b_GR_ridehail * GR +           

    # Specific individual characteristics
    b_rider_ridehail * rider +      
      
    # TT (Travel Time) indirect effects
    (b_stud_TT_ridehail * STUD + 
     b_ridfreq_TT_ridehail * ridfreq
     ) * Rid_TT +
  
    # AT (Access Time) indirect effects
    (
     b_stud_AT_ridehail * STUD
     ) * Rid_AT
  )

  
  ### Define settings for MNL model component
  mnl_settings = list(
    alternatives = c(walk=1,scooty=2, expbike=3, ondemand=4, ridehail=5),
    avail        = list(walk=av_walk,
                        scooty=av_Sc, 
                        expbike=av_Bik, 
                        ondemand=av_OnD, 
                        ridehail=av_Rid ),
    choiceVar    = choice,
    utilities    = V
  )
  
  ### Compute probabilities using MNL model
  P[["model"]] = apollo_mnl(mnl_settings, functionality)
  
  ### Take product across observation for same individual
  P = apollo_panelProd(P, apollo_inputs, functionality)
  
  ### Average across inter-individual draws
  P = apollo_avgInterDraws(P, apollo_inputs, functionality)
  
  ### Prepare and return outputs of function
  P = apollo_prepareProb(P, apollo_inputs, functionality)
  return(P)
}

```

## Prepare useful functions

Apollo estimation result cleaning

```{r result_cleaning_function}

sign_coef_LLC = function(modelname='model_name', mat) {
  
  # add significance star
  
  add_sign <- function(p_value) {
    if (p_value < 0.01) {
      return("***")
    } else if (p_value < 0.05) {
      return("**")
    } else if (p_value < 0.1) {
      return("*")
    } else {
      return("")
    }
  }
  
  copy_mat = mat
  
  mat[,1] = format(mat[,1], nsmall = 3)
  mat[,1] <- paste(mat[,1], sapply(mat[,3], add_sign), sep = "")
  mat = data.frame(mat[,1])
  mat = rbind(mat, model$nIndivs, model$nObs, model$nFreeParams,
              round(model$LLC,2), round(model$LLout,2), 
              round(model$adjRho2_0,2), round(model$BIC,2))
  
  SE = c(abs(copy_mat[,2]), rep(NA, 7))
  
  mat = cbind(mat, SE)

  rownames(mat)[(nrow(mat)-6):nrow(mat)] = c('nIndivs','nobs','nbeta','LL(c)',
                          'LL(beta)','Adj.rho2','BIC')
  colnames(mat) = c(modelname,'s.e.')

  return(mat)
}

```

Save cleaned results to xlsx:

- create new xlsx

- if xlsx exists, append

- if sheet exists, rewrite

```{r SaveToXlsx}

save_to_xlsx <- function(data, file_name, sheet_name) {
  if (file.exists(file_name)) {

    wb <- loadWorkbook(file_name)
    
    if (sheet_name %in% names(wb)) {
      removeWorksheet(wb, sheet_name)
    }
    
    addWorksheet(wb, sheet_name)
    writeData(wb, sheet_name, data, rowNames = TRUE)
    saveWorkbook(wb, file_name, overwrite = TRUE)
    
  } else {
    wb <- createWorkbook()
    addWorksheet(wb, sheet_name)
    writeData(wb, sheet_name, data, rowNames = TRUE)
    saveWorkbook(wb, file_name, overwrite = TRUE)
  }
}

```


## Estimation and clean result

```{r estimation}

# MODEL ESTIMATION 

model = apollo_estimate(apollo_beta, apollo_fixed, apollo_probabilities, apollo_inputs,
                        estimate_settings = list(silent = FALSE,
                                                 maxIterations=400,
                                                 writeIter = F))

# FORMATTED OUTPUT (TO SCREEN)

modelOutput_settings = list() #controls for summary
modelOutput_settings[['printPVal']] = 2 #include p-values and robust standard errros

est_mat = as.matrix(round(apollo_modelOutput(model, modelOutput_settings)[,c(1,5,7)],3),ncol=3)

est_mat = est_mat[complete.cases(est_mat), ]

est_table = data.frame(sign_coef_LLC(model_name, est_mat))

# Save model to memory

assign(paste0('apollo_',model_name),model)
assign(paste0('estmat_',model_name),est_mat)
assign(paste0('esttable_',model_name),est_table)

# print summarized model report to console

print.data.frame(get(paste0('esttable_',model_name)))

# save model to file

apollo_saveOutput(get(paste0('apollo_',model_name)),saveOutput_settings = list(printPVal=2))

# save summarized model report to file

save_to_xlsx(get(paste0('esttable_',model_name)),
             paste0("Summarized_",
                    "addmixed",
                    ".xlsx"),
             model_name
             )

# output_list <- list(model_name = get(paste0('esttable_',model_name))
#                     )
# 
# write.xlsx(output_list, 
#            file = paste0("Summarized_",
#                          model_name,
#                          ".xlsx"),
#            rowNames = T, colNames = T,
#            overwrite = FALSE,
#            append = TRUE)

# load from file

# model = apollo_loadModel(get(paste0('apollo_',model_name)))

```

## Post-estimation

```{r}

#Unconditionals#

# unconditional = apollo_unconditionals(get(paste0('apollo_',model_name)),
#                                       apollo_probabilities, 
#                                       apollo_inputs)
# 
# #Plotting the random parameter distribution
# 
# tiff("atranpar.tiff", width =6, height = 4, units = 'in', res = 500)
# 
# #par(mar=c(4,6,1,0))
# 
# plot(density(unconditional$b_at_sc), xlim=c(-1,1), ylim=c(0,7),xlab="", 
#      main="", col="red", lwd=2)
# lines(density(unconditional$b_at_bik), col= "orange",lwd=2)
# lines(density(unconditional$b_at_OnD), col= "blueviolet",lwd=2)
# lines(density(unconditional$b_at_Rid), col= "blue",lwd=2)
# legend("topright", inset = .04, legend=c("e-scooter","bike-share",
#       "ondemand transit","ride-hail"), col=c( "red", "orange",
#       "blueviolet", "blue"), lty = 1, lwd = 2, box.lty = 0)
# 
# dev.off()

```


