---
title: "FMLM_tt_at_Cor"
author: "Kenny Chandra Wijaya"
date: "`r Sys.Date()`"
header-includes:
  - \usepackage{fontspec} # use fontspec package
  - \usepackage{xeCJK}    # use xeCJK package
  - \usepackage{setspace}\setstretch{1.2}
  - \usepackage{titlesec}
  # - \newcommand{\sectionbreak}{\clearpage}
  - \usepackage{siunitx}
  - \newcolumntype{d}{S[table-format=3.3]}
output: 
  pdf_document: 
    pandoc_args: ["--lua-filter=color-text.lua"]
    latex_engine: xelatex
    number_sections: yes
---

```{r setup, include=FALSE}
# Overal all chunk settings

knitr::opts_chunk$set(echo = FALSE,knitr.kable.NA = ''
                      )

# I/O
library(readxl)
library(openxlsx)
library(writexl)
# rmd render
library(kableExtra)
library(pander)
library(extrafont)
library(showtext)
# Data cleaning
library(tidyverse)
library(janitor)
# Math
library(scales)
library(MASS)
library(Matrix)
# Plotting
library(ggpubr)
library(ggplot2)
library(gridExtra)
library(grid)
# DCM
library(apollo)
```

# Preparations

## load data

```{r Initialization}

# clean the environment

rm(list = ls())

database<- read.csv("C:\Users\kenny\OneDrive - purdue.edu\Documents\Kenny's File\Transportation Literature\Fall 2024\Prasanna\Papers\FMLM 3-05-24\Purdue\fmlm_5.csv")

database <- database %>% 
  mutate(Pri_car = ifelse(as.numeric(Q3.1_2)==1,1,0)) %>% 
  mutate(Pri_tra = ifelse(as.numeric(Q3.1_2) %in% c(4,5),1,0))
database <- database %>% dplyr::filter(av_car==0)

```

## load apollo setting

```{r}

# apollo Initialization

model_name = 'addmixed4'
Drawmethod = "sobol"
Ndraws = 2000
apollo_initialise()

# Set core controls

apollo_control = list(
  modelName       = model_name,
  modelDescr      = "MIXL with balanced sociodemo + latent, no correlations",
  indivID         = "cid", 
  outputDirectory = "output",
  nCores          = parallel::detectCores()-1,
  seed            = 123
)

# Define choice and data

choiceAnalysis_settings <- list(
  alternatives = c(walk=1,scooty=2, expbike=3, ondemand=4, ridehail=5),
  avail        = list(walk=database$av_walk, scooty= database$av_Sc, expbike=database$av_Bik,
                      ondemand=database$av_OnD,ridehail=database$av_Rid),
  choiceVar    = database$choice,
  explanators  = database[,c("STUD","Gend")],
  rows = database$choice!=0
)


### Vector of parameters

apollo_beta = c(
  # ASC (Alternative Specific Constants)
  asc_scooty         = 0,
  asc_expbike        = 0,
  asc_ondemand       = 0,
  asc_ridehail       = 0,
  
  # Travel time coefficients (mu and sigma)
  b_tt_walk_mu       = -1,
  b_tt_scooty_mu     = -1,
  b_tt_expbike_mu    = -1,
  b_tt_OnD_mu        = -1,
  b_tt_Rid_mu        = -1,
  
  b_tt_walk_sig      = 1,
  b_tt_scooty_sig    = 1,
  b_tt_expbike_sig   = 1,
  b_tt_OnD_sig       = 1,
  b_tt_Rid_sig       = 1,

  # Access time coefficients (mu and sigma)
  b_at_scooty_mu     = -1,
  b_at_expbike_mu    = -1,
  b_at_OnD_mu        = -1,
  b_at_Rid_mu        = -1,
  
  b_at_scooty_sig    = 1,
  b_at_expbike_sig   = 1,
  b_at_OnD_sig       = 1,
  b_at_Rid_sig       = 1,

  # Cost coefficient
  b_cost             = -1,

  # Travel Time (TT) indirect effects
  b_stud_TT_ondemand    = 0,
  b_stud_TT_ridehail    = 0,

  b_scfreq_TT_scooty    = 0,

  b_ridfreq_TT_ridehail = 0,

  # Access Time (AT) indirect effects

  b_stud_AT_ondemand    = 0,
  b_stud_AT_ridehail    = 0,
  
  b_bikfreq_AT_expbike  = 0,
  
  # STUD coefficients for on-demand ride and ride-hailing modes
  b_stud_ondemand    = 0,

  # Latent variable coefficient
  b_GR_ridehail      = 0,  
  
  # Specific individual characteristics
  b_rider_ondemand   = 0,
  b_rider_ridehail   = 0,

  b_scfreq_scooty    = 0,

  b_scbik_plane_scooty  = 0,
  b_scbik_plane_expbike = 0
  )



apollo_fixed = c()

### Set parameters for generating draws

apollo_draws = list(
  interDrawsType = Drawmethod,
  interNDraws = Ndraws,
  interNormDraws= c(# Travel time
                    "draws_tt_sc",
                    "draws_tt_OnD",
                    "draws_tt_bik",
                    "draws_tt_Rid",
                    "draws_tt_walk",
                    "draws_tt_car",
                    # Access time
                    "draws_at_sc","draws_at_OnD",
                    "draws_at_bik","draws_at_Rid")
)

### Create random parameters

apollo_randCoeff = function(apollo_beta, apollo_inputs){
  randcoeff = list()
  
  randcoeff[["b_tt_walk"]] = b_tt_walk_mu + b_tt_walk_sig * draws_tt_walk
  
  randcoeff[["b_tt_sc"]]   = b_tt_scooty_mu + b_tt_scooty_sig * draws_tt_sc  
    
  randcoeff[["b_tt_bik"]]  = b_tt_expbike_mu + b_tt_expbike_sig * draws_tt_bik
                            
  randcoeff[["b_tt_OnD"]]  = b_tt_OnD_mu + b_tt_OnD_sig * draws_tt_OnD
                             
  randcoeff[["b_tt_Rid"]]  = b_tt_Rid_mu + b_tt_Rid_sig * draws_tt_Rid
                            
  randcoeff[["b_at_sc"]]   = b_at_scooty_mu + b_at_scooty_sig * draws_at_sc
                             
  randcoeff[["b_at_bik"]]  = b_at_expbike_mu + b_at_expbike_sig * draws_at_bik
  
  randcoeff[["b_at_OnD"]]  = b_at_OnD_mu + b_at_OnD_sig * draws_at_OnD
     
  randcoeff[["b_at_Rid"]]  = b_at_Rid_mu + b_at_Rid_sig * draws_at_Rid
  
  return(randcoeff)
}

apollo_inputs = apollo_validateInputs()

apollo_probabilities=function(apollo_beta, apollo_inputs, functionality="estimate"){
  
  ### Attach inputs and detach after function exit
  apollo_attach(apollo_beta, apollo_inputs)
  on.exit(apollo_detach(apollo_beta, apollo_inputs))
  
  ### Create list of probabilities P
  P = list()
  
  ### List of utilities: these must use the same names as in mnl_settings, order is irrelevant
  
  V = list()
  
  # Walk is the base alternative, no individual characteristics included
  V[["walk"]] = (
    b_tt_walk * Walk_TT
  )
  
  # Utility function for electric scooter mode
  V[["scooty"]] = (
    asc_scooty + 
    b_tt_sc * Sc_TT + 
    b_at_sc * Sc_AT + 
    b_cost * Sc_Cost +
      
    # Specific individual characteristics
    b_scfreq_scooty * scfreq +     
    b_scbik_plane_scooty * scbik_plane +
    
    # TT (Travel Time) indirect effects
    (
     b_scfreq_TT_scooty * scfreq 
     ) * Sc_TT 
  
    # AT (Access Time) indirect effects

  )
  
  # Utility function for bike-sharing mode
  V[["expbike"]] = (
    asc_expbike + 
    b_tt_bik * Bik_TT + 
    b_at_bik * Bik_AT + 
    b_cost * Bik_Cost +
  
    # Specific individual characteristics
    b_scbik_plane_expbike * scbik_plane +
    
    # TT (Travel Time) indirect effects
  
    # AT (Access Time) indirect effects
    (
     b_bikfreq_AT_expbike * bikfreq
     ) * Bik_AT
  )
  
  # Utility function for on-demand ride mode
  V[["ondemand"]] = (
    asc_ondemand + 
    b_tt_OnD * OnD_TT + 
    b_at_OnD * OnD_AT + 
    b_cost * OnD_Cost +
  
    # General individual characteristics (GIC)
    b_stud_ondemand * STUD +       

    # Specific individual characteristics
    b_rider_ondemand * rider +      
      
    # TT (Travel Time) indirect effects
    (
     b_stud_TT_ondemand * STUD 
     ) * OnD_TT +
  
    # AT (Access Time) indirect effects
    (
     b_stud_AT_ondemand * STUD 
     ) * OnD_AT
  )
  
  # Utility function for ride-hailing mode
  V[["ridehail"]] = (
    asc_ridehail + 
    b_tt_Rid * Rid_TT + 
    b_at_Rid * Rid_AT + 
    b_cost * Rid_Cost +
  
    # General individual characteristics (GIC)
    b_GR_ridehail * GR +           

    # Specific individual characteristics
    b_rider_ridehail * rider +      
      
    # TT (Travel Time) indirect effects
    (b_stud_TT_ridehail * STUD + 
     b_ridfreq_TT_ridehail * ridfreq
     ) * Rid_TT +
  
    # AT (Access Time) indirect effects
    (
     b_stud_AT_ridehail * STUD
     ) * Rid_AT
  )

  
  ### Define settings for MNL model component
  mnl_settings = list(
    alternatives = c(walk=1,scooty=2, expbike=3, ondemand=4, ridehail=5),
    avail        = list(walk=av_walk,
                        scooty=av_Sc, 
                        expbike=av_Bik, 
                        ondemand=av_OnD, 
                        ridehail=av_Rid ),
    choiceVar    = choice,
    utilities    = V
  )
  
  ### Compute probabilities using MNL model
  P[["model"]] = apollo_mnl(mnl_settings, functionality)
  
  ### Take product across observation for same individual
  P = apollo_panelProd(P, apollo_inputs, functionality)
  
  ### Average across inter-individual draws
  P = apollo_avgInterDraws(P, apollo_inputs, functionality)
  
  ### Prepare and return outputs of function
  P = apollo_prepareProb(P, apollo_inputs, functionality)
  return(P)
}

model = apollo_loadModel('addmixed4_OLD1')

apollo_modelOutput(model)
```




# Post-estimation

## variable Categories

Behavioral Data

- Walk_TT: "Travel time for walking mode (minutes)"

- Sc_TT: "Travel time for electric scooter (minutes)"

- Bik_TT: "Travel time for bike-sharing (minutes)"

- OnD_TT: "Travel time for on-demand ride (minutes)"

- Rid_TT: "Travel time for ride-hailing mode (minutes)"

- Sc_AT: "Access time for electric scooter (minutes)"

- Bik_AT: "Access time for bike-sharing (minutes)"

- OnD_AT: "Access time for on-demand ride (minutes)"

- Rid_AT: "Access time for ride-hailing mode (minutes)"

- Sc_Cost: "Cost for electric scooter trip"

- Bik_Cost: "Cost for bike-sharing trip"

- OnD_Cost: "Cost for on-demand ride trip"

- Rid_Cost: "Cost for ride-hailing mode trip"

Personal Characteristic Data 12

- STUD: "Student status (1 = Full-time or part-time student, 0 = Non-student)"

- Gend: "Gender (1 = Male, 0 = Female or other)"

- Emp: "Employment status (1 = Full-time employment, 0 = Other)"

- Pola: "Political inclination (1 = Liberal or leaning liberal, 0 = Other)"

- Pri_car: "Private car ownership status (1 = Owns a car, 0 = Does not own)"

- GR: "Latent variable (Pro-environmental behavior)"

- Pri_tra: "Primary transportation is public transport (1 = Public transport, 0 = Other)"

- rider: "Bus rider status (1 = Bus user, 0 = Non-bus user)"

- scfreq: "Electric scooter usage frequency (1 = At least 1-2 times per week, 0 = Less frequent)"

- bikfreq: "Bike-sharing usage frequency (1 = At least 1-2 times per week, 0 = Less frequent)"

- ridfreq: "Ride-hailing usage frequency (1 = At least 1-2 times per week, 0 = Less frequent)"

- scbik_plane: "Planned variable for combined bike-sharing and electric scooter use (1 = Likely to use more when more protected lanes are available)"

## clean database

```{r}
# Define the required fields and their order
ordered_fields <- c( 'cid',
  "choice", "Walk_TT", "Sc_TT", "Bik_TT", "OnD_TT", "Rid_TT",
  "Sc_AT", "Bik_AT", "OnD_AT", "Rid_AT",
  "Sc_Cost", "Bik_Cost", "OnD_Cost", "Rid_Cost",
  "Gend", "Pri_car", 'Pri_tra',"scfreq", "bikfreq", "Emp", "STUD", "Pola", 
  "rider", "ridfreq", "scbik_plane", "GR", "av_walk", "av_Sc", 
  "av_Bik", "av_OnD", "av_Rid"
)

# Filter and reorder columns in the database based on ordered_fields
database <- database[, ordered_fields[ordered_fields %in% names(database)], drop=FALSE]

# Display the structure of the filtered and ordered database
glimpse(database)

data = database

```

## divide data by sociodemographic controls

```{r}

vars <- c("scfreq", "bikfreq", "STUD", "rider", "ridfreq", "scbik_plane")

temp <- data.frame()

for (var in vars) {
  temp_var <- database %>%
    tabyl(!!sym(var)) %>%
    mutate(Varname = var) %>%             # Add Varname column
    rename(Value = !!sym(var)) %>%
    mutate(percent = percent(n / sum(n)))  # Format percent to two decimal places   
  
  # Reorder columns by creating a new dataframe structure
  temp_var <- temp_var[, c("Varname", "Value", "n", "percent")] 
  
  temp <- rbind(temp, temp_var)
}

# Individual level statistics
temp_individual <- data.frame()

for (var in vars) {
  temp_var <- database %>%
    group_by(cid) %>%                       # Group by cid (individuals)
    summarize(Value = first(!!sym(var))) %>%  # Get the first value per individual for the variable
    ungroup() %>%
    tabyl(Value) %>%                        # Tabulate individual level
    mutate(Varname = var) %>%
    rename(nindiv = n) %>%                  # Rename n to nindiv (count of individuals)
    mutate(percent = percent(nindiv / sum(nindiv), accuracy = 0.01))  # Format percent
  
  # Reorder columns
  temp_var <- temp_var[, c("Varname", "Value", "nindiv", "percent")]
  
  temp_individual <- rbind(temp_individual, temp_var)
}

social_demo_observe = cbind(temp, temp_individual[,c(3,4)])

print.data.frame(social_demo_observe)  
```

vars <- c("scfreq", "bikfreq", "STUD", "rider", "ridfreq", "scbik_plane")

```{r}

# Function to split and name data subsets based on provided criteria
split_data <- function(data, var_name, value) {
  # Create a subset based on the variable and value
  subset_data <- subset(data, data[[var_name]] == value)
  
  # Dynamically name the subset
  assign(paste0("dt_", var_name, "_", value), subset_data, envir = .GlobalEnv)
}

# Split database by scfreq (0 and 1)
split_data(database, "scfreq", 0)
split_data(database, "scfreq", 1)

# Split database by bikfreq (0 and 1)
split_data(database, "bikfreq", 0)
split_data(database, "bikfreq", 1)

# Split database by STUD
split_data(database, "STUD", 0)
split_data(database, "STUD", 1)

# Split database by rider
split_data(database, "rider", 0)
split_data(database, "rider", 1)

# Split database by ridfreq (0 and 1)
split_data(database, "ridfreq", 0)
split_data(database, "ridfreq", 1)

# Split database by scbik_plane (0 and 1)
split_data(database, "scbik_plane", 0)
split_data(database, "scbik_plane", 1)
```

## 1. Unconditionals

```{r}
database = data

unconditional = apollo_unconditionals(model,
                                      apollo_probabilities,
                                      apollo_inputs)

# Access time plot with appropriate title
plot(density(unconditional$b_at_sc), xlim=c(-1,1), ylim=c(0,2.5), xlab="",
     main="Access Time's Unconditional Distribution", col="red", lwd=2)
lines(density(unconditional$b_at_bik), col= "orange",lwd=2)
lines(density(unconditional$b_at_OnD), col= "pink",lwd=2)
lines(density(unconditional$b_at_Rid), col= "blue",lwd=2)
legend("topleft", inset = .02, legend=c("e-scooter","bike-share",
      "ondemand transit","ride-hail"), col=c( "red", "orange",
      "pink", "blue"), lty = 1, lwd = 2, box.lty = 0)

# Travel time plot with appropriate title
plot(density(unconditional$b_tt_sc), xlim=NULL, ylim=c(0,1.2), xlab="",
     main="Travel Time's Unconditional Distribution", col="red", lwd=2)
lines(density(unconditional$b_tt_bik), col= "orange",lwd=2)
lines(density(unconditional$b_tt_OnD), col= "pink",lwd=2)
lines(density(unconditional$b_tt_Rid), col= "blue",lwd=2)
lines(density(unconditional$b_tt_walk), col= "green",lwd=2)
legend("topleft", inset = .02, legend=c("e-scooter","bike-share",
      "ondemand transit","ride-hail",'walk'), col=c( "red", "orange",
      "pink", "blue",'green'), lty = 1, lwd = 2, box.lty = 0)

```


## 2. Conditionals by groups

### prepare function

```{r}


plot_con_bygroup <- function(ap_model, ap_prob, ap_inputs, VarName,
                             at_xlim = NULL, at_ylim=NULL,
                             tt_xlim = NULL, tt_ylim=NULL) {
  
  # Compute conditional distributions
  conditional <- apollo_conditionals(ap_model, ap_prob, ap_inputs)
  nindiv <- length(conditional[["b_tt_walk"]][["ID"]])
  
  # Prepare data for access time coefficients plot
  access_data <- data.frame(
    x = c(conditional[["b_at_sc"]][["post.mean"]],
          conditional[["b_at_bik"]][["post.mean"]],
          conditional[["b_at_OnD"]][["post.mean"]],
          conditional[["b_at_Rid"]][["post.mean"]]),
    group = factor(rep(c("e-scooter", "bike-share", "ondemand", "ride-hail"), 
                       times = c(length(conditional[["b_at_sc"]][["post.mean"]]), 
                                 length(conditional[["b_at_bik"]][["post.mean"]]),
                                 length(conditional[["b_at_OnD"]][["post.mean"]]), 
                                 length(conditional[["b_at_Rid"]][["post.mean"]]))))
  )
  
  # Generate access time plot using ggplot2
  plot1 <- ggplot(access_data, aes(x = x, color = group)) +
    geom_density(size = 1) +
    labs(title = paste("Density Plot of Access Time Coefficients (", VarName, ")\n nindiv =", nindiv, sep=""),
         x = "", y = "Density") +
    xlim(at_xlim) +
    ylim(at_ylim) +
    theme_minimal() +
    theme(
      #text = element_text(family = "Times New Roman"),  # Set font to Times New Roman
      plot.title = element_text(hjust = 0.5, size = 14),  # Center title and set font size
      legend.title = element_blank(),  # Remove legend title
      legend.position = "top"  # Move legend to top
    ) +
    guides(color = guide_legend(override.aes = list(linetype = 1))) +  # Change legend to lines
    scale_color_manual(values = c("red", "orange", "pink", "blue"))

  # Prepare data for travel time coefficients plot
  travel_data <- data.frame(
    x = c(conditional[["b_tt_sc"]][["post.mean"]],
          conditional[["b_tt_bik"]][["post.mean"]],
          conditional[["b_tt_OnD"]][["post.mean"]],
          conditional[["b_tt_Rid"]][["post.mean"]],
          conditional[["b_tt_walk"]][["post.mean"]]),
    group = factor(rep(c("e-scooter", "bike-share", "ondemand", "ride-hail", "walk"), 
                       times = c(length(conditional[["b_tt_sc"]][["post.mean"]]), 
                                 length(conditional[["b_tt_bik"]][["post.mean"]]),
                                 length(conditional[["b_tt_OnD"]][["post.mean"]]), 
                                 length(conditional[["b_tt_Rid"]][["post.mean"]]),
                                 length(conditional[["b_tt_walk"]][["post.mean"]]))))
  )
  
  # Generate travel time plot using ggplot2
  plot2 <- ggplot(travel_data, aes(x = x, color = group)) +
    geom_density(size = 1) +
    labs(title = paste("Density Plot of Travel Time Coefficients (", VarName, ")\n nindiv =", nindiv, sep=""),
         x = "", y = "Density") +
    xlim(tt_xlim) +
    ylim(tt_ylim) +
    theme_minimal() +
    theme(
      #text = element_text(family = "Times New Roman"),  # Set font to Times New Roman
      plot.title = element_text(hjust = 0.5, size = 14),  # Center title and set font size
      legend.title = element_blank(),  # Remove legend title
      legend.position = "top"  # Move legend to top
    ) +
    guides(color = guide_legend(override.aes = list(linetype = 1, shape = NA))) +  # Change legend to lines
    scale_color_manual(values = c("red", "orange", "pink", "blue", "green"))
  
  # Dynamically assign the plots to global environment
  assign(paste0('p_', "dt_", VarName, '_at'), plot1, envir = .GlobalEnv)
  assign(paste0('p_', "dt_", VarName, '_tt'), plot2, envir = .GlobalEnv)
}

```

### plotting

vars <- c("scfreq", "bikfreq", "STUD", "rider", "ridfreq", "scbik_plane")

#### (1) scfreq

```{r}
Varname = "scfreq_0"
database = get(paste0("dt_",Varname))
apollo_inputs = apollo_validateInputs()
plot_con_bygroup(model, apollo_probabilities, apollo_inputs, Varname,
                 c(-0.5,0.25),c(0,10),c(-4,1),c(0,2))
get(paste0('p_', "dt_", Varname, '_at'))
get(paste0('p_', "dt_", Varname, '_tt'))

Varname = "scfreq_1"
database = get(paste0("dt_",Varname))
apollo_inputs = apollo_validateInputs()
plot_con_bygroup(model, apollo_probabilities, apollo_inputs, Varname,
                 c(-0.5,0.25),c(0,10),c(-4,1),c(0,2))
get(paste0('p_', "dt_", Varname, '_at'))
get(paste0('p_', "dt_", Varname, '_tt'))

```


#### (2) bikfreq

vars <- c("scfreq", "bikfreq", "STUD", "rider", "ridfreq", "scbik_plane")

```{r}
Varname = "bikfreq_0"
database = get(paste0("dt_",Varname))
apollo_inputs = apollo_validateInputs()
plot_con_bygroup(model, apollo_probabilities, apollo_inputs, Varname,
                 c(-0.5,0.25),c(0,10),c(-4,1),c(0,2))
get(paste0('p_', "dt_", Varname, '_at'))
get(paste0('p_', "dt_", Varname, '_tt'))

Varname = "bikfreq_1"
database = get(paste0("dt_",Varname))
apollo_inputs = apollo_validateInputs()
plot_con_bygroup(model, apollo_probabilities, apollo_inputs, Varname,
                 c(-0.5,0.25),c(0,10),c(-4,1),c(0,2))
get(paste0('p_', "dt_", Varname, '_at'))
get(paste0('p_', "dt_", Varname, '_tt'))

```

#### (3) STUD

vars <- c("scfreq", "bikfreq", "STUD", "rider", "ridfreq", "scbik_plane")

```{r}
Varname = "STUD_0"
database = get(paste0("dt_",Varname))
apollo_inputs = apollo_validateInputs()
plot_con_bygroup(model, apollo_probabilities, apollo_inputs, Varname,
                 c(-0.5,0.25),c(0,10),c(-4,1),c(0,2))
get(paste0('p_', "dt_", Varname, '_at'))
get(paste0('p_', "dt_", Varname, '_tt'))

Varname = "STUD_1"
database = get(paste0("dt_",Varname))
apollo_inputs = apollo_validateInputs()
plot_con_bygroup(model, apollo_probabilities, apollo_inputs, Varname,
                 c(-0.5,0.25),c(0,10),c(-4,1),c(0,2))
get(paste0('p_', "dt_", Varname, '_at'))
get(paste0('p_', "dt_", Varname, '_tt'))

```

#### (4) rider

vars <- c("scfreq", "bikfreq", "STUD", "rider", "ridfreq", "scbik_plane")

```{r}
Varname = "rider_0"
database = get(paste0("dt_",Varname))
apollo_inputs = apollo_validateInputs()
plot_con_bygroup(model, apollo_probabilities, apollo_inputs, Varname,
                 c(-0.5,0.25),c(0,10),c(-4,1),c(0,2))
get(paste0('p_', "dt_", Varname, '_at'))
get(paste0('p_', "dt_", Varname, '_tt'))

Varname = "rider_1"
database = get(paste0("dt_",Varname))
apollo_inputs = apollo_validateInputs()
plot_con_bygroup(model, apollo_probabilities, apollo_inputs, Varname,
                 c(-0.5,0.25),c(0,10),c(-4,1),c(0,2))
get(paste0('p_', "dt_", Varname, '_at'))
get(paste0('p_', "dt_", Varname, '_tt'))
```

#### (5) ridfreq

vars <- c("scfreq", "bikfreq", "STUD", "rider", "ridfreq", "scbik_plane")

```{r}
Varname = "ridfreq_0"
database = get(paste0("dt_",Varname))
apollo_inputs = apollo_validateInputs()
plot_con_bygroup(model, apollo_probabilities, apollo_inputs, Varname,
                 c(-0.5,0.25),c(0,10),c(-4,1),c(0,2))
get(paste0('p_', "dt_", Varname, '_at'))
get(paste0('p_', "dt_", Varname, '_tt'))

Varname = "ridfreq_1"
database = get(paste0("dt_",Varname))
apollo_inputs = apollo_validateInputs()
plot_con_bygroup(model, apollo_probabilities, apollo_inputs, Varname,
                 c(-0.5,0.25),c(0,10),c(-4,1),c(0,2))
get(paste0('p_', "dt_", Varname, '_at'))
get(paste0('p_', "dt_", Varname, '_tt'))
```

#### (6) scbik_plane

vars <- c("scfreq", "bikfreq", "STUD", "rider", "ridfreq", "scbik_plane")

```{r}
Varname = "scbik_plane_0"
database = get(paste0("dt_",Varname))
apollo_inputs = apollo_validateInputs()
plot_con_bygroup(model, apollo_probabilities, apollo_inputs, Varname,
                 c(-0.5,0.25),c(0,10),c(-4,1),c(0,2))
get(paste0('p_', "dt_", Varname, '_at'))
get(paste0('p_', "dt_", Varname, '_tt'))

Varname = "scbik_plane_1"
database = get(paste0("dt_",Varname))
apollo_inputs = apollo_validateInputs()
plot_con_bygroup(model, apollo_probabilities, apollo_inputs, Varname,
                 c(-0.5,0.25),c(0,10),c(-4,1),c(0,2))
get(paste0('p_', "dt_", Varname, '_at'))
get(paste0('p_', "dt_", Varname, '_tt'))
```

### prepare combine plot functions

```{r}
# Define the packaged function
generate_combined_plot <- function(Varname) {
  
  # Dynamically retrieve the plot objects by name
  p_dt_at_0 <- get(paste0("p_dt_", Varname, "_0_at"))
  p_dt_at_1 <- get(paste0("p_dt_", Varname, "_1_at"))
  p_dt_tt_0 <- get(paste0("p_dt_", Varname, "_0_tt"))
  p_dt_tt_1 <- get(paste0("p_dt_", Varname, "_1_tt"))

  # Modify the plot subtitles and remove individual legends
  p_dt_at_0 <- p_dt_at_0 + ggtitle(paste0("Access time (", Varname, "=0)")) + theme(legend.position = "none")
  p_dt_at_1 <- p_dt_at_1 + ggtitle(paste0("Access time (", Varname, "=1)")) + theme(legend.position = "none")
  p_dt_tt_0 <- p_dt_tt_0 + ggtitle(paste0("Travel time (", Varname, "=0)")) + theme(legend.position = "none")
  p_dt_tt_1 <- p_dt_tt_1 + ggtitle(paste0("Travel time (", Varname, "=1)")) + theme(legend.position = "none")

  # Extract the unified legend
  legend_dt <- get_legend(p_dt_tt_1 + 
                          theme(legend.position = "top", legend.direction = "horizontal"))

  # Combine the four plots into a 2x2 layout
  combined_plot_dt <- grid.arrange(
    arrangeGrob(p_dt_at_0, p_dt_at_1, p_dt_tt_0, p_dt_tt_1, 
                ncol = 2, nrow = 2),
    ncol = 1
  )

  # Generate the final layout with a main title, legend, and combined plots
  p_comb_con_Varname <- grid.arrange(
    textGrob(paste0("Coefficients Distributions by Subgroup (", Varname, ")"), 
             gp = gpar(fontsize = 16
                       #, fontfamily = "Times New Roman"
                       )), 
    legend_dt,  
    combined_plot_dt, 
    ncol = 1, heights = c(0.1, 0.1, 1)  # Adjust height ratios
  )

  # Use assign to save the plot with a dynamic variable name
  assign(paste0("p_comb_con_", Varname), p_comb_con_Varname, envir = .GlobalEnv)
  
  # Return the generated plot object
  return(p_comb_con_Varname)
}
```



### check combined plots

vars <- c("scfreq", "bikfreq", "STUD", "rider", "ridfreq", "scbik_plane")

```{r}
generate_combined_plot("scfreq")
generate_combined_plot("bikfreq")
generate_combined_plot("STUD")
generate_combined_plot("rider")
generate_combined_plot("ridfreq")
generate_combined_plot("scbik_plane")

```

## 3. VOT

```{r}

database = data

deltaMethod_settings = list(
  expression = c(
    VOT_tt_walk    = "b_tt_walk_mu/b_cost",
    VOT_tt_scooty  = "b_tt_scooty_mu/b_cost",
    VOT_tt_expbike = "b_tt_expbike_mu/b_cost",
    VOT_tt_OnD     = "b_tt_OnD_mu/b_cost",
    VOT_tt_Rid     = "b_tt_Rid_mu/b_cost",
    VOT_at_scooty  = "b_at_scooty_mu/b_cost",
    VOT_at_expbike = "b_at_expbike_mu/b_cost",
    VOT_at_OnD     = "b_at_OnD_mu/b_cost",
    VOT_at_Rid     = "b_at_Rid_mu/b_cost"
  ),
  varcov = 'robust'  # include heteroscedasticity
)

se_apollo = apollo_deltaMethod(model, deltaMethod_settings)
# se_apollo$`t-ratio (0)`
# se_apollo$`Rob t-ratio (0)`
se_apollo$Significance <- cut(se_apollo$`Rob t-ratio (0)`,
                              breaks = c(-Inf, 1.64, 1.96, 2.58, Inf),
                              labels = c("", "*", "**", "***"))

se_apollo$Value <- paste0(se_apollo$Value, se_apollo$Significance)

se_apollo_simplified <- se_apollo[, c("Expression", "Value")]

print.data.frame(se_apollo_simplified)

```

## 4. Marginal Effects

```{r}
# Set database
database = data
apollo_inputs = apollo_validateInputs(database = database)

# Predict baseline probabilities using the model and apollo_probabilities
P_predicted <- apollo_prediction(model, apollo_probabilities, apollo_inputs)

# Initialize an empty data frame to store results
results <- data.frame(Data = character(), 
                      walk = numeric(), 
                      scooty = numeric(),
                      expbike = numeric(),
                      ondemand = numeric(),
                      ridehail = numeric(),
                      stringsAsFactors = FALSE)

# Define the list of variables (not parameters) for which we want to calculate marginal effects
param_list <- c(
  "Walk_TT",
  "Sc_TT",
  "Bik_TT",
  "OnD_TT",
  "Rid_TT",
  "Sc_AT",
  "Bik_AT",
  "OnD_AT",
  "Rid_AT"
)

# Loop over each variable and calculate marginal effects
for (param_name in param_list) {
  
  # Save the original values of the current variable
  original_values <- database[[param_name]]
  
  # Increase the values of the current variable by 0.1% (small change)
  database[[param_name]] <- original_values * 1.0001
  
  # Revalidate apollo_inputs after modifying the data
  apollo_inputs_new <- apollo_validateInputs(database = database)
  
  # Predict new probabilities with the updated variable values
  P_predicted_new <- apollo_prediction(model, apollo_probabilities, apollo_inputs_new)
  
  # Calculate marginal effects for each choice alternative
  me_walk <- (P_predicted_new$walk - P_predicted$walk) / (original_values * 0.0001)
  me_scooty <- (P_predicted_new$scooty - P_predicted$scooty) / (original_values * 0.0001)
  me_expbike <- (P_predicted_new$expbike - P_predicted$expbike) / (original_values * 0.0001)
  me_ondemand <- (P_predicted_new$ondemand - P_predicted$ondemand) / (original_values * 0.0001)
  me_ridehail <- (P_predicted_new$ridehail - P_predicted$ridehail) / (original_values * 0.0001)
  
  # Compute the average marginal effect across all individuals
  mean_me_walk <- mean(me_walk, na.rm = TRUE)
  mean_me_scooty <- mean(me_scooty, na.rm = TRUE)
  mean_me_expbike <- mean(me_expbike, na.rm = TRUE)
  mean_me_ondemand <- mean(me_ondemand, na.rm = TRUE)
  mean_me_ridehail <- mean(me_ridehail, na.rm = TRUE)
  
  # Append the result to the results data frame
  results <- rbind(results, data.frame(Data = param_name, 
                                       walk = mean_me_walk, 
                                       scooty = mean_me_scooty, 
                                       expbike = mean_me_expbike, 
                                       ondemand = mean_me_ondemand, 
                                       ridehail = mean_me_ridehail))
  
  # Restore the original values of the variable for the next iteration
  database[[param_name]] <- original_values
}

# Format the marginal effects as percentages with two decimal places
results$walk <- percent(results$walk, accuracy = 0.01)
results$scooty <- percent(results$scooty, accuracy = 0.01)
results$expbike <- percent(results$expbike, accuracy = 0.01)
results$ondemand <- percent(results$ondemand, accuracy = 0.01)
results$ridehail <- percent(results$ridehail, accuracy = 0.01)

# Display the final results
print.data.frame(results)


```

Check data distribution
```{r}
vars <- list(
  Walk_TT = data$Walk_TT,
  Sc_TT = data$Sc_TT,
  Bik_TT = data$Bik_TT,
  OnD_TT = data$OnD_TT,
  Rid_TT = data$Rid_TT,
  Sc_AT = data$Sc_AT,
  Bik_AT = data$Bik_AT,
  OnD_AT = data$OnD_AT,
  Rid_AT = data$Rid_AT
)

result <- lapply(names(vars), function(name) {
  cat("\nTable for", name, ":\n")
  print(table(vars[[name]]))
  cat("\n-------------------------------\n")
})


```








